### 2014-06-10 Studies/2012_Chicago/Report/Chicago_Fire.tex
Chicago_Fire: update report and add figures

A firefighter inside is susceptible to second degree burn injuries when exposed to temperatures 55 $^{\circ}$C (130 $^{\circ}$F) \cite{designation2003c}. While firefighters wear protective gear, the gear only offers finite levels of protection. The polycarbonate in a self contained breathing apparatus begins to soften when the material temperature reaches approximately 140 $^{\circ}$C (284 $^{\circ}$F) \cite{mensch2011emergency}. Structural fire fighting coats and pants are tested to withstand temperatures of 260 $^{\circ}$C (500 $^{\circ}$F) \cite{nfpa2013standard}. Extended exposure to elevated temperatures can result in significant heat being transferred to the firefighter, putting him/her at risk. For the analysis of the model predicted exposure temperatures in the flow path, we have set the maximum visualized temperature to be all values that are 260 $^{\circ}$C (500 $^{\circ}$F) or above. This means that temperatures visualized in the highest range in the model output are at a minimum of 260 $^{\circ}$C (500 $^{\circ}$F); the temperatures could be higher.

\section{Tactical Considerations}

\chapter{Summary}

>>>>

A firefighter is susceptible to second degree burn injuries when exposed to temperatures 55 $^{\circ}$C (130 $^{\circ}$F) \cite{designation2003c}. While firefighters wear protective gear, gear only offers finite levels of protection. The polycarbonate in a self contained breathing apparatus begins to soften when the material temperature reaches approximately 140 $^{\circ}$C (284 $^{\circ}$F) \cite{mensch2011emergency}. Structural fire fighting coats and pants are tested to withstand temperatures of 260 $^{\circ}$C (500 $^{\circ}$F) \cite{nfpa2013standard}. Extended exposure to elevated temperatures can result in significant heat transferred to the firefighter, putting him/her at risk. Therefore, for analysis of the model predicted exposure temperatures, we set the maximum visualized temperature to be all values that are 260 $^{\circ}$C (500 $^{\circ}$F) or above. This means that temperatures visualized in the highest range in the model output are at a minimum of 260 $^{\circ}$C (500 $^{\circ}$F); predicted temperatures could in fact be higher.

Prior to door failure, the elevated temperatures on the second floor were limited to the rear porch, Fig.~(\ref{fig:temp_top_159}). Temperatures are highest at the Beta-Charlie corner where there is sufficient oxygen due to self-ventilation. Five seconds after the door fails, hot gases have traveled the length of the interior hallway and approached the doorway into the kitchen, approximately 4.5 m (14.8 ft) away from the source fire (cf Fig.~(\ref{fig:temp_top_165})). One minute after the door fails, the hallway temperatures are nominally at or above 260 $^{\circ}$C (500 $^{\circ}$F) through out the hallways corridor (cf Fig.~(\ref{fig:temp_top_220})). This illustrates the the high temperature gases seen five seconds after failure were not just a burst but rather the front of a sustained exhaust gas flow from the fire burning on the rear porch. The conditions in this hallway changed from tenable to high-hazard very rapidly following the door failure. Fig.~(\ref{fig:temp_top_220}) also shows that that temperatures in the kitchen dropped below 100 $^{\circ}$C (212 $^{\circ}$F) as the large supply of air cooled/diluted the exhaust gases. Based on post-incident images of the kitchen and second floor stairwell, these predicted temperatures are realistic.

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/west_50th_baseline_top_159}
\caption{PLACEHOLDER: Temperature contours on the second floor, 1 second prior to door failure, 6ft above the floor.}
\label{fig:temp_top_159}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/west_50th_baseline_top_165}
\caption{PLACEHOLDER: Temperature contour on the second floor, 5 seconds after door failure, 6ft above the floor.}
\label{fig:temp_top_165}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/west_50th_baseline_top_200}
\caption{PLACEHOLDER: Temperature contour on the second floor, 60 seconds after door failure, 6ft above the floor.}
\label{fig:temp_top_220}
\end{figure}

\section{Tactical Considerations}
In the last fifteen years, there have been fires throughout the country where the change in a flow path has had a significant impact on firefighter and occupant safety:

{\bf UNDER DEVELOPMENT}

\begin{itemize}
\item NE Washington D.C., 1999
\item Keokuk, IA, 1999
\item Houston, TX, 2011
\item Riverdale Heights, MD, 2012 
\end{itemize}

\chapter{Summary}
NIST's Fire Dynamics Simulator was used to provide insight into the fire dynamics within a 2 1/2 story single family structure in Chicago, Il that resulted in the death of a firefighter. The modeled fuel, fire size, and fire growth rate were estimated taking into account all available information: NIOSH post-incident report~\cite{NIOSH:Bowyer}, post-incident pictures, videos, and relevant literature. This resulted in a specified source fire of 12.3 MW, however based on ventilation the model predicted a peak HRR of $\approx$ 10 MW. The fire began in the attic and spread into the rear (Charlie side) porch. Spread of high temperature, toxic gases into the structure was limited by a closed steel door on the second floor. After prolonged exposure to elevated temperature and pressure, the door eventually failed. Prior to the door failure, a FF and Captain were on the second floor in the interior hallway. The door failure lead to the creation of flow path towards the front (Alpha side) door on the first floor of the structure. The flow path was comprised of toxic, high temperature combustion gases that rapidly changed the conditions where the FF and Captain were working. The rapid change in hazard level overcame the Captain before he could find egress to safety. After a MayDay, the Captain was rescued from the structure but later succumbed to his injuries.
